---
title: Genus-level ASCI
output: 
  html_document
self_contained: yes
---

```{r, echo = F}
knitr::opts_chunk$set(message = F, warning = F)
```

### Genus traits table

* The existing traits table at the species (or higher) level is formatted to show the genus designation, assigned FinalID (species or higher), trait category, and estimate for the trait.

```{r, warning = F, message = F}
library(tidyverse)
library(ASCI)

data(taxain_spp)

# add a complete genus column to the traits lookup table
# select only the trait categories used in selected metrics
x <- pmmilkup$traits %>%
  dplyr::select(Phylum, Class, Order, Family, Genus1, FinalIDassigned, everything()) %>% 
  dplyr::select(-orig.FinalID, -AlgaeList) %>% 
  mutate_if(is.factor, as.character) %>% 
  rowwise() %>% 
  mutate(Genus1 = do({
    
    # use genus1 if not empty
    if(Genus1 != '')
      return(Genus1)
    
    chk <- c(Phylum, Class, Order, Family, Genus1)
    
    # get first of FinalIDassigned if all higher taxa cats are empty
    if(all(chk == '')){
      
      out <- strsplit(FinalIDassigned, ' ')[[1]][1]
      
      # get last non-empty cat  
    } else {
      
      out <- chk[chk != ''] %>% rev %>% .[1]
      
    }
    
    return(out)
    
  })
  ) %>% 
  dplyr::select(-Phylum, -Class, -Order, -Family) %>% 
  unique %>% 
  gather('trait', 'est', -Genus1, -FinalIDassigned) %>% 
  rename(FinalID = FinalIDassigned) %>% 
  arrange(Genus1, FinalID, trait)

print(tbl_df(x), n = 30)
```

* The observed species data is joined with the formatted traits table by FinalID to get counts of species within each genus, i.e., which species in a genus are more common in the dataset
```{r}
# join species level data with genus traits to summarize counts by data dist
to_summ <- taxain_spp %>% 
  left_join(x, by = 'FinalID') %>% 
  mutate(
    Genus1 = ifelse(is.na(Genus1), gsub(' .*$', '', FinalID), Genus1),
    est = ifelse(est %in% '', NA, est)
  ) %>% 
  arrange(Genus1, FinalID, trait) %>% 
  select(SampleID, Genus1, FinalID, trait, est)

print(tbl_df(to_summ), n = 30)
```

* The joined data are summarized to count the number of observations for each species in a genus
```{r}
# summarize trait counts by data dist
traits_gen <- to_summ %>% 
  group_by(Genus1, trait, est) %>% 
  summarise(n = length(est)) %>% 
  group_by(Genus1, trait) %>% 
  mutate(prp = n / sum(n))

print(tbl_df(traits_gen), n = 30)
```

* Traits are assigned to the genus level if more than 75% of species in the genus have a given trait, otherwise NA.  Assignments at higher taxonomic resolution (e.g., family) remain the same.
```{r}
traits_gen <- traits_gen %>% 
  group_by(Genus1, trait) %>% 
  filter(prp > 0.75) %>% 
  ungroup %>% 
  dplyr::select(-prp, -n) %>% 
  spread(trait, est) %>% 
  rename(FinalIDassigned = Genus1)

knitr::kable(traits_gen[1:30, ])
```

### Indicator genus analysis

* Multipatt function (indicspecies package) requires site categories as a vector and species abundances as a matrix, row.names match for both
* Sites in development dataset grouped by 'rc', 'int', 'str'
```{r, eval = T}
library(tidyverse)
library(indicspecies)
library(ASCI)

data(taxain_gen)
data(sitcat)

##
# groups for indic
grpin <- sitcat %>% 
  filter(cls %in% c('int', 'rc', 'str')) %>% 
  mutate(
    cls = fct_recode(cls, Intermediate = 'int',  Reference = 'rc', Stressed = 'str'),
    cls = as.character(cls)
  ) %>% 
  dplyr::select(-psa) %>% 
  spread(SampleID, cls) %>% 
  unlist

print(tbl_df(grpin), n = 30)
```

* Abundance of genera at sites (in long format)
```{r}
##
# filter taxonomy data by sites in dev dataset

# format abundance ests
abuin <- taxain_gen %>% 
  dplyr::select(SampleID, FinalID, BAResult, Result) %>% 
  mutate(
    abund = as.numeric(pmax(BAResult, Result, na.rm = T))
  ) %>% # combine abundance 
  dplyr::select(-BAResult, -Result) %>% 
  group_by(SampleID, FinalID) %>% 
  summarise(abund = sum(abund, na.rm = T)) %>%
  ungroup

print(tbl_df(abuin), n = 30)
```

```{r}
# make wide format, location to rowid
abuin <- abuin %>% 
  spread(FinalID, abund, fill = 0) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('SampleID')

##
# make sure sites match between grpin and abuin, order is the same
grpin <- grpin[names(grpin) %in% rownames(abuin)]
abuin <- abuin %>% 
  .[rownames(.) %in% names(grpin), ] %>% 
  .[match(rownames(.), names(grpin)), ]

```

* Run indicator species analysis, output include:
  * the A value describes the affinity of a species for each group, this sums to one for each species across the groups, it indicates which group a species is most likely to belong
  * the B value describes the proportion of sites within a group that have a species, if found in the group
  * the final stat value for indic analysis considers A and B
  * multipatt returns a pvalue for each species
  * the species indicators can be further selected by filtering with an A or B threshold
  * the coverage function describes the proportion of sites in each group where one or more indicators are found for a given threshold, with A = 0 being the default
  * using a more conservative, larger A will remove indicator species but this can be done up to a threshold until coverage begins to decline

```{r}
##
# run indicator analysis
res <- multipatt(abuin, grpin, duleg = T, control = how(nperm = 999))

head(res$A)
head(res$B)
head(res$sign)
```

* Significant indicator species
```{r}
# table of indicator species by pvalue
indic <- res$sign %>%
  rownames_to_column('FinalID') %>% 
  filter(p.value < 0.05) %>% 
  mutate(designation = factor(index, 
                              levels = c('1', '2', '3'), 
                              labels = c('Intermediate', 'Reference', 'Stressed'))
         
  ) %>% 
  select(FinalID, designation, stat, p.value)

dim(indic)
head(indic)
```

* Using a conservative selection of indicator species
```{r}
# all a values
aval <- res$A %>% 
  data.frame %>% 
  rownames_to_column('FinalID') %>% 
  gather('designation', 'aval', -FinalID)

# find maximum a values without decreasing coverage
chks <- seq(0, 1, by = 0.005)
toplo <- sapply(chks, function(At) coverage(x = abuin, y = res, At = At)) %>% 
  t %>% 
  data.frame %>% 
  mutate(chks = chks) %>% 
  gather('designation', 'cvg', -chks)
athr <- toplo %>% 
  group_by(designation) %>%
  filter(cvg == max(cvg)) %>% 
  filter(chks == max(chks)) %>% 
  select(-cvg) %>% 
  ungroup
ggplot(toplo, aes(x = chks, y = cvg, group = designation, colour = designation)) +
  geom_vline(data = athr, aes(xintercept = chks, colour = designation), linetype = 'dashed') +
  scale_x_continuous('At') + 
  geom_line() +
  theme_bw()

# join indic with a val, a threshold
indic <- indic %>% 
  left_join(aval, by = c('FinalID', 'designation')) %>% 
  left_join(athr, by = c('designation')) %>% 
  filter(aval >= chks)

head(indic)

# indicators <- indic %>% 
#   dplyr::select(FinalID, designation) %>% 
#   rename(FinalIDassigned = FinalID)
# save(indicators, file = 'C:/Users/Marcus.SCCWRP2K/Desktop/indicators.RData', compress = 'xz')
```

### Raw metrics

All metrics were estimated for the genus-level data using the new traits table and indicator genera.
```{r}
data(rawmet_gen)
data(rawmet_spp)

dim(rawmet_gen)
head(rawmet_gen)
unique(rawmet_gen$met)
```

