# README

```{r, message = F, warning = F, echo = F}
library(knitr)
library(tidyverse)
library(ASCI)
opts_chunk$set(message = F, warning = F)

data(pkgdat)
data(orgdat)
data(sitcat)
```

Materials for evaluating the ASCI index. See the sister repository [here](https://github.com/SCCWRP/ASCI) for the complete package.

Files in `data` folder created in `R/dat_proc.R` unless otherwise noted:

* `orgdat.RData` Original ASCI results from Susie's code
* `pkgdat.RData` Complete ASCI results from `sitein` and `taxain`, using ASCI package
* `sitcat.RData` Site categories as reference (cal/val), intermediate, and stressed
* `sitein.RData` Processed site data from raw data
* `taxain.RData` Processed taxonomy data from raw data

```{r, fig.height = 6, fig.width = 8}

pkgdat <- pkgdat %>% 
  scores %>% 
  dplyr::select(SampleID, taxa, OoverE, MMI) %>% 
  rename(
    oe = OoverE, 
    mmi = MMI,
    grp = taxa
    ) %>% 
  gather('ind', 'scr', oe, mmi) %>% 
  mutate(dat = 'pkg') %>% 
  dplyr::select(SampleID, ind, grp, scr, dat)

toplo <- rbind(pkgdat, orgdat) %>% 
  spread(dat, scr)

ggplot(toplo, aes(x = org, y = pkg)) +
  facet_wrap(ind ~ grp) +
  geom_point(alpha = 0.6) + 
  geom_abline(intercept = 0, slope = 1)
```

Summary of performance:

```{r}



```



